---
title: "Nursing Homes"
author: "Eye on Ohio, Savanna Strott writer, Lucia Walinchus Editor"
date: "7/23/2021"
output: html_document
---

```{r call packages}
library(dplyr)
library(ggplot2)
library("RSocrata")
library(lubridate)
library(tidyverse) 
library(DescTools)
library(psych)
library(fastDummies)
library(purrr)
library(dkanr)


```

## Nurse Staffing

This comes from the Centers for Medicare and Medicaid Services [Payroll-Based Journal Daily Nurse Staffing.](https://data.cms.gov/quality-of-care/payroll-based-journal-daily-nurse-staffing)

# set up R work space

options(stringsAsFactors = FALSE, scripen = 999)

# set working directory

save.image("OhioNH.RData")
load("OhioNH.RData")


# import data

"Total"=read.csv("NH6.27.21.csv")

#Separate for only NH that submitted data
DataUS = Total[Total$Submitted.Data == "Y",]

```{r}
#How many does that exclude?
nrow(Total)-nrow(DataUS)
#What percentage is that total?
(nrow(Total)-nrow(DataUS))/nrow(Total)
```



#Separate for only NH that submitted data in Ohio
OhioData = DataUS[DataUS$Provider.State == "OH",]

###Look at shortages for key weeks (first, peak, latest)

```{r}
#Find shortage for the first week, 05/24/20, for Ohio
OH52420=OhioData[OhioData$Week.Ending == "05/24/20",]
length(OH52420$Shortage.of.Aides[OH52420$Shortage.of.Aides == "Y"])
nrow(OH52420)
```

```{r}
#Find shortage for the first week, 05/24/20, for the US
US52420 = DataUS[DataUS$Week.Ending == "05/24/20",]
length(US52420$Shortage.of.Aides[US52420$Shortage.of.Aides =="Y"])
nrow(US52420)
```

```{r}
#Find shortage for 12/13/20 (Ohio's covid cases peak) for Ohio
OH121320 = OhioData[OhioData$Week.Ending =="12/13/20",]
length(OH121320$Shortage.of.Aides[OH121320$Shortage.of.Aides == "Y"])
nrow(OH121320)
```

```{r}
#Look at US 12/13/20
US121320= DataUS[DataUS$Week.Ending == "12/13/20",]
length(US121320$Shortage.of.Aides[US121320$Shortage.of.Aides == "Y"])
nrow(US121320)
```

```{r}
#Look at Ohio for 6/6/21 (give the NHs/data a two-ish week window to get caught up-Applebaum)
#Update with most current numbers (but 2-3 weeks delayed) pre-publication
OH060621 = OhioData[OhioData$Week.Ending =="06/06/21",]
length(OH060621$Shortage.of.Aides[OH060621$Shortage.of.Aides == "Y"])
nrow(OH060621)
```

```{r}
#Look at US for 6/6/21
US060621= DataUS[DataUS$Week.Ending == "06/06/21",]
length(US060621$Shortage.of.Aides[US060621$Shortage.of.Aides == "Y"])
nrow(US060621)
```



### Rank states by shortages
```{r}
#Remove unnecessary columns, focusing on staffing shortages
USsmall = DataUS[-7:-65]
USsmall = USsmall[-10:-220]

#Separate files by year by separating out the month, day, year

USsmall = transform(USsmall,Date = as.character(Week.Ending))
USsplit = strsplit(USsmall$Week.Ending, split = "/")
USsmall = transform(USsmall, 
                    Date1=sapply(USsplit,"[[",1), 
                    Date2=sapply(USsplit, "[[",2),
                    Date3=sapply(USsplit, "[[",3))
```



```{r}
#Make vectors for necessary info for ranking, based on rows from 2020
aides20 = USsmall$Shortage.of.Aides[grep("20",USsmall$Date3)]
nurse20 = USsmall$Shortage.of.Nursing.Staff[grep("20",USsmall$Date3)]
month20= USsmall$Date1[grep("20",USsmall$Date3)]
state20= USsmall$Provider.State[grep("20",USsmall$Date3)]
name20 = USsmall$Provider.Name[grep("20",USsmall$Date3)]

#Create the 2020 data frame
Y20=data.frame(Provider.Name=name20,
               Provider.State=state20,
               Month=month20,
               Shortage.of.Aides=aides20,
               Shortage.of.Nurses=nurse20)

#Make dummy columns so that the binary (y/n) is numeric and separated

dumY20 = dummy_cols(Y20, select_columns=c("Shortage.of.Aides"))

```


```{r}
#Group by staff and find the percentage of shortage

FinalY20 = dumY20 %>% group_by(Provider.State) %>%
  summarise(Count_shortages=sum(Shortage.of.Aides_Y),
            Percentage_shortages=
              Count_shortages/
              sum(Shortage.of.Aides_Y,Shortage.of.Aides_N)*100)

FinalY20=FinalY20[order(FinalY20$Percentage_shortages, decreasing=TRUE),]
```


```{r}
ggplot(FinalY20, aes(x=reorder(Provider.State,-Count_shortages),Count_shortages))+
  geom_col()+
  theme(axis.text.x = element_text(angle = 45, vjust = 1.2, hjust = 1.1,size = 8))
```

```{r}
ggplot(FinalY20, aes(x=reorder(Provider.State,-Percentage_shortages),Percentage_shortages))+
  geom_col()+
  theme(axis.text.x = element_text(angle = 45, vjust = 1.2, hjust = 1.1,size = 8))
```



write.csv(FinalY20, "FinalY20.csv", fileEncoding = "UTF-8")

#Do the same for 2021

```{r}
aides21 = USsmall$Shortage.of.Aides[grep("21",USsmall$Date3)]
nurse21 = USsmall$Shortage.of.Nursing.Staff[grep("21",USsmall$Date3)]
month21= USsmall$Date1[grep("21",USsmall$Date3)]
state21= USsmall$Provider.State[grep("21",USsmall$Date3)]
name21 = USsmall$Provider.Name[grep("21",USsmall$Date3)]

Y21=data.frame(Provider.Name=name21,
               Provider.State=state21,
               Month=month21,
               Shortage.of.Aides=aides21,
               Shortage.of.Nurses=nurse21)

dumY21 = dummy_cols(Y21, select_columns=c("Shortage.of.Aides"))


FinalY21 = dumY21 %>% group_by(Provider.State) %>%
  summarise(Count_shortages=sum(Shortage.of.Aides_Y),
            Percentage_shortages=
              Count_shortages/
              sum(Shortage.of.Aides_Y,Shortage.of.Aides_N)*100)

FinalY21=FinalY21[order(FinalY21$Percentage_shortages, decreasing=TRUE),]
```


```{r}
ggplot(FinalY21, aes(x=reorder(Provider.State,-Count_shortages),Count_shortages))+
  geom_col()+
  theme(axis.text.x = element_text(angle = 45, vjust = 1.2, hjust = 1.1,size = 8))
```


```{r}
ggplot(FinalY21, aes(x=reorder(Provider.State,-Percentage_shortages),Percentage_shortages))+
  geom_col()+
  theme(axis.text.x = element_text(angle = 45, vjust = 1.2, hjust = 1.1,size = 8))
```

write.csv(FinalY21, "FinalY21.csv", fileEncoding = "UTF-8")

#Covid Nursing home Data 

Information on COVID-19 reported by nursing homes to the [CDC’s National Healthcare Safety Network (NHSN) COVID-19 Long Term Care Facility Module.](https://data.cms.gov/covid-19/covid-19-nursing-home-data)





